#   Reproducible Research Course Project-Vidushi Bhakat
===============================
## Step 1  
### Code for reading/processing the data set
```{r, cache=T,echo=TRUE} 
unzip("repdata_data_activity.zip")
activitydata <- read.csv("activity.csv")
head(activitydata)
```
## Step 2   
### Histogram of total number of steps taken everyday

```{r,echo=T,cache=T}
aggregateactivity <- tapply(activitydata$steps,activitydata$date,sum)
head(aggregateactivity)
activitytotal <- data.frame(day.number=1:61,totalactivity=aggregateactivity)
hist(activitytotal$totalactivity,col="green", breaks =15 ,xlab="Steps",ylab="Frequency",main="Total Steps Histogram")
```

## Step 3  
### Mean and median of total steps taken everyday
```{r,echo=T,cache=T}
mean(activitytotal$totalactivity,na.rm=T)
median(activitytotal$totalactivity,na.rm=T)
```



##  Step 4  
###  Time series plot of average steps taken 
 ```{r,echo=T,cache=T}
 library(ggplot2)
 intervalactivity <- aggregate(steps~interval,data=activitydata, mean, na.action=na.omit)
 h <- ggplot(intervalactivity,aes(interval,steps))
 h + geom_line(col="blue")+ggtitle("Average steps per time interval")+
 xlab("Interval") + ylab("Steps")
 ```



##   Step 5   
###   5 min interval with maximum number of steps
 ```{r,echo=T,cache=T}
 library(dplyr)
 intervalactivity %>% select(interval,steps) %>% filter(steps ==max(intervalactivity$steps))
```

## Step 6  
### Imputing missing data 
1. Total number of missing values
```{r, echo=T, cache=TRUE}
sum(is.na(activitydata$steps))
```

2. Replacing Missing Values
```{r,echo=T, cache= TRUE}
activitydata$completesteps <- ifelse(is.na(activitydata$steps), round(intervalactivity$steps[match(activitydata$interval, intervalactivity$interval)],0), activitydata$steps)
```

3. New data set with missing values filled in
```{r,echo=T,cache=TRUE}
completeactivity <- data.frame(steps=activitydata$completesteps,interval=activitydata$interval,date=activitydata$date)
head(completeactivity)
```


## Step 7
### Histogram of daily total steps with missing data filled in
```{r,echo=TRUE,cache=TRUE}
aggregatecompleteactivity <- tapply(completeactivity$steps,completeactivity$date, sum)
hist(aggregatecompleteactivity,col="green",breaks=20,xlab="Steps",ylab="Frequency",main="Total Steps Histogram")
```


### Calculating mean and median of total daily activity 
```{r,echo=T,cache=T}
mean(aggregatecompleteactivity)
median(aggregatecompleteactivity)
```
Imputing missing data have only a little and transcurable impact on the mean and the median of the total daily number of steps.


## Step 8
### Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

1.Creating a new factor variable with two levels weekend and weekday
```{r,echo=T,cache=T}
completeactivity$realdate <- as.Date(completeactivity$date, format = "%Y-%m-%d")
completeactivity$weekday <- weekdays(completeactivity$realdate)
completeactivity$daytype <- ifelse(completeactivity$weekday=='Saturday' | completeactivity$weekday=='Sunday', 'weekend','weekday')
head(completeactivity)
 
```

2. 5 min time series plot
```{r,echo=T,cache=T}
intervalactivityDT <- aggregate(steps~interval+daytype,data=completeactivity, FUN=mean, na.action=na.omit)
j <- ggplot(intervalactivityDT,aes(interval,steps))
j +  geom_line(col="blue") + facet_grid(.~daytype) + ggtitle(" Average steps per 5 min time interval- weekday v/s weekends") + xlab("Interval") + ylab("Steps")
```

 
 
 
 
 